---
title: "TCV Vaccination Status"
author: "Chris LeBoa"
date: "8/19/2021"
output: word_document
---

## Evaluating Vaccination Status within the Navi Mumbai Project 

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# Add libraries 
library(tidyverse)
library(haven)
library(lubridate)
library(fs)
library(tableone)




ccs_data_ch_cdc <- read_sas("/Users/ChrisLeBoa/Box Sync/CDC_Typhoid_data/ccs_ch_2020_03_18.sas7bdat")
 ##This is the Child based file with the addition of specific CDC variables 

data_location <- "/Users/ChrisLeBoa/Box Sync/CDC_Typhoid_data/tcv_all__2021_04_18_vax.sas7bdat"
 ##dataset with all hospital data in it up to april 2021

#Read in data 
all_data <- read_sas(data_location)

```

#IVI wanted an estimate of the % of Typoid cases that were hospitalized

We ask about three day fever with a recall period but hospitalization in terms of a 12 month recall period. 

```{r, scientific = FALSE}
options(scipen=999) #This turns off scientific notation

data_for_ivi <- 
ccs_data_ch_cdc %>% 
  mutate(
    seek_care_hospital = if_else( ccs_trtloc_a == "1" | ccs_trtloc_b == "1" | ccs_trtloc_c == "1" |ccs_trtloc_d == "1"| ccs_trtloc_e == "1" | ccs_trtloc_f == "1" | ccs_trtloc_g == "1" | ccs_trtloc_i == "1", 1, 0), 
    across(where(is.numeric), ~na_if(., 99))) %>% 
  group_by(PHASE) %>% 
  summarise(
    total_fever_per_month = sum(fev3day, na.rm = TRUE), 
    total_fever_per_year = total_fever_per_month * (365 / 30), 
    total_children = n(), 
    prop_children_fev_mon = total_fever_per_month / total_children, 
    prop_children_fev_year = total_fever_per_year / total_children,
    hospitalized_year = sum(hosp, na.rm = TRUE),
    percent_fever_hospitalized_year = 
      sum(hosp, na.rm = TRUE) / total_fever_per_year * 100, 
    percent_fever_sought_care_one_month = 
      sum(seek_care_hospital, na.rm = TRUE) / total_fever_per_month * 100
  ) %>% 
  pivot_longer(-PHASE, names_to = "Variable") %>% 
  pivot_wider(names_from = PHASE, values_from = value) %>%
  mutate_if(is.numeric, ~round(., 3))

data_for_ivi

data_for_ivi %>% 
  write_csv("/Users/ChrisLeBoa/GitHub/typhoid_research/india_typhoid/data/Data_for_IVI_220723.csv")


```

## IVI also wanted us to get a 95% confidence interval for these numbers above 

Calculating a 95% confidence interval from a binomial is using the equation 
p  +/-  z*(√p(1-p) / n)

For a 95% confidence interval use z = 1.96
```{r}
n1 <- 5378
f1 <- 437
h1 <- 195
fy1 <- 5317
p1 <- f1 / n1
ph1 <- h1 / fy1

n2 <- 4446
f2 <- 361
h2 <- 142
fy2 <-  4392
p2 <- f2 / n2
ph2 <- h2 / fy2

z <- 1.96

# ci 1 
p1
p1 - z * (sqrt(p1 * (1 - p1)) / n1)
p1 + z * (sqrt(p1 * (1 - p1)) / n1)

#ci hosph 1
ph1
ph1 - z * (sqrt(ph1 * (1 - ph1)) / fy1)
ph1 + z * (sqrt(ph1 * (1 - ph1)) / fy1)

# ci 2 
p2
p2 - z * (sqrt(p2 * (2 - p2)) / n2)
p2 + z * (sqrt(p2 * (1 - p2)) / n2)

#ci hosph 2
ph2
ph2 - z * (sqrt(ph2 * (2 - ph2)) / fy2)
ph2 + z * (sqrt(ph2 * (2 - ph2)) / fy2)


```
```{r}

```


